---
title: "Assignment - Bayesian Networks"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---
-------------------------------------------------------------------------------


## Loading the Data
```{r}
data_assignment_3 <- read.csv(file = "C:/Users/Suma Marri/Documents/GitHub/COVID/Cleaned-Data.csv",
                              colClasses = "character"
                          )
data_assignment_3
```


# Preparing the dataset for bnlearn package to have all factor variables
Removing Country and making sure we have factor variables for Bayesian networks
```{r}
d <- subset(data_assignment_3, select = -c(Country))
summary(d)

for(j in 1:ncol(d)){  
  d[,j] <- factor(as.numeric(d[,j]))}
summary(d)
```

# Building the Bayesian Network Models
Constructing the Bayesian Models using a constraint-based, score-based, hybrid, and a local discovery algorithm. Using the the algorithms benchmarked by Dr. Smith with regards to computation time.
```{r}
require(bnlearn)
d_algorithms <- c("iamb.fdr", "hc", "h2pc", "aracne")
list_bnlearn <- list()
for(j in d_algorithms) try({
  list_bnlearn[[j]] <- do.call(
    what = j,
    args = list(x = d)
    )
  }
  )
list_bnlearn
```
The score-based algorithm hc() and the hybrid algorithm h2pc() produce directed graphs.
Lets see if we can produce directed graphs with any other constraint-based and local discovery algorithms
```{r}
d_algorithms <- c("pc.stable", "gs", "iamb", "inter.iamb", "mmpc", "si.hiton.pc", "hpc", "chow.liu")
list2_bnlearn <- list()
for(j in d_algorithms) try({
  list2_bnlearn[[j]] <- do.call(
    what = j,
    args = list(x = d)
    )
  },silent = TRUE
  )
list2_bnlearn
```

## bic
```{r}
d_algorithms <- c("iamb.fdr", "hc", "h2pc", "aracne")

M_score_bic <- list()
for (j in d_algorithms) try({
  M_score_bic[j] <- score(
    x=list_bnlearn[[j]],
    data = d,
    type = "bic"
  )
}
)  
M_score_bic <- data.frame(M_score_bic)
M_score_bic
```
## aic
```{r}
d_algorithms <- c("iamb.fdr", "hc", "h2pc", "aracne")

M_score_aic <- list()
for (j in d_algorithms) try({
  M_score_aic[j] <- score(
    x=list_bnlearn[[j]],
    data = d,
    type = "aic"
  )
}
)  
M_score_aic <- data.frame(M_score_aic)
M_score_aic
```
## loglik
```{r}
d_algorithms <- c("iamb.fdr", "hc", "h2pc", "aracne")

M_score_loglik <- list()
for (j in d_algorithms) try({
  M_score_loglik[j] <- score(
    x=list_bnlearn[[j]],
    data = d,
    type = "loglik"
  )
}
)  
M_score_loglik <- data.frame(M_score_loglik)
M_score_loglik
```
## bde
```{r}
d_algorithms <- c("iamb.fdr", "hc", "h2pc", "aracne")

M_score_bde <- list()
for (j in d_algorithms) try({
  M_score_bde[j] <- score(
    x=list_bnlearn[[j]],
    data = d,
    type = "bde"
  )
}
)  
M_score_bde <- data.frame(M_score_bde)
M_score_bde
```
## bds
```{r}
d_algorithms <- c("iamb.fdr", "hc", "h2pc", "aracne")

M_score_bds <- list()
for (j in d_algorithms) try({
  M_score_bds[j] <- score(
    x=list_bnlearn[[j]],
    data = d,
    type = "bds"
  )
}
)  
M_score_bds <- data.frame(M_score_bds)
M_score_bds
```


## Model Score Comparison
```{r}
#Combining the scores in a table df
g <- rbind(M_score_bic,M_score_aic,M_score_loglik, M_score_bde, M_score_bds)
rownames(g) <- c("bic", "aic", "loglik", "bde", "bds")
g
#Formatting the table to see which algorithm performs better
h <- data.frame(t(g))
colnames(h) <- rownames(g)
h
#Sorting largest to smallest in terms of performance of algorithms
sorted_h <- h[order(h$bic,h$aic,h$loglik,h$bde,h$bds),]
sorted_h
```
```{r}
apply(sorted_h, 2, FUN=max)
```
# Visualizing the model hc()
## Setting the node and edge attributes
```{r}
#Node Attributes
hc_covid <- hc(d)
v_nodes <- nodes(hc_covid)
names(v_nodes) <- v_nodes
strength_covid <- arc.strength( 
  x = hc_covid,  
  data = d
)
n_nodes <- nnodes(hc_covid)

v_fillcolor <- viridis::viridis(n_nodes)
names(v_fillcolor) <- v_nodes

v_shape <- c(
  rep("circle",floor(n_nodes/3)),
  rep("ellipse",floor(n_nodes/3)),
  rep("box",n_nodes - 2*floor(n_nodes/3))
  )
names(v_shape) <- v_nodes

#Edge Attributes
v_edges <- paste0(strength_covid[,"from"],"~",
                  strength_covid[,"to"])
names(v_edges) <- v_edges

v_edgecolor <- v_fillcolor[strength_covid[,"from"]]
names(v_edgecolor) <- v_edges
```
## Plotting the model
```{r}
#Converting the bnlearn model to a graphNEL model
# library(BiocManager)
# install.packages("BiocManager")
# BiocManager::install(c("graph", "RBGL", "Rgraphviz"))
graphNEL_covid <- as.graphNEL(hc_covid)
```

```{r}
Rgraphviz::plot(  
  x = graphNEL_covid, 
  y = "twopi",  attrs = list(),   
  nodeAttrs = list(    
    fillcolor = v_fillcolor,    
    shape = v_shape  
    ),   
  edgeAttrs = list(    
    label = v_edges,    
    weight = graph::edgeWeights(graphNEL_covid),    
    color = v_edgecolor,    
    fontcolor = v_edgecolor
    )
  )
```

# Predict the target variable

## Predicting the target
```{r}
fit_model <- bn.fit(
  x = hc_covid,
  data = d
)
pred_table <- data.frame(pred = predict(fit_model, node = "Nasal.Congestion", data = d),actual = d$Nasal.Congestion)
summary(pred_table)

```

## Evaluating Model Fit
```{r}
#Calculating AUC
AUC_model <- Metrics::auc(pred_table$actual,pred_table$pred)
print(paste0("AUC:", AUC_model))
#Calculating Model Accuracy
Accuracy_model <- Metrics::accuracy(pred_table$actual,pred_table$pred)
print(paste0("Accuracy: ", Accuracy_model))
# Constructing the confusion matrix
# remove.packages("rlang")
# install.packages("rlang")
# install.packages("vctrs")
# install.packages("pillar")

confusion_matrix_model <- caret::confusionMatrix(pred_table$pred,pred_table$actual,"1")

cm <- data.frame(confusion_matrix_model$byClass)
cm
```

```{r}
# Repeated 2-fold Cross-validation
cv_model <- bn.cv(
  data = d,
  bn = "hc",
  k = 2,
  runs = 2
)
cv_model
```



